---
title: "Installing EpiModel"
author: "Network Modeling for Epidemics"
output:
  html_document:
    css: style.css
    fig_width: 8
    highlight: kate
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA)
```

These are instructions for installing `EpiModel` for use in the workshop, *Network Modeling for Epidemics*.

To follow the workshop tutorials, please have `R`, `EpiModel`, and its dependent packages installed before the workshop using the instructions below. If you have any questions about software installation, please contact the course instructor, Samuel Jenness at sjenness@uw.edu.


### R and Rstudio
`EpiModel` runs on the `R` statistical software platform. The current version of `R` is 3.1.1; this version should be installed on your computer. It is available on [CRAN](http://cran.r-project.org/).

The instructors will use `Rstudio` for the workshop. This is a popular front-end IDE (integrated development environment) for `R`, and is recommended if you are relatively new to `R`. It is available on the [Rstudio website](http://www.rstudio.com/).


### EpiModel
With an open `R` console window, install `EpiModel` and its main dependencies by typing the following.
```{r, eval=FALSE}
install.packages("EpiModel")
```
`R` may prompt you to select a mirror site. In addition to `EpiModel` itself, this step will also install or update the needed **statnet** software, including the `ergm`, `tergm`, `network`, and `networkDynamic` packages.


Some visualization examples in the workshop use two packages not automatically installed with `EpiModel`, which may be installed as follows:
```{r, eval=FALSE}
install.packages(c("ndtv", "shiny"))
```

After installation, test that `EpiModel` can load correctly by typing:
```{r, message=FALSE}
library(EpiModel)
```

Finally, check that the installed versions of `EpiModel`, `ergm`, `tergm`, `network`, and `networkDynamic` all match those listed below under "other attached packages":
```{r}
sessionInfo()
```

<br>

### Help and Tutorials
After loading the software, feel free to browse the help file documentation and tutorials in advance of the workshop. This is particularly recommended if you are new to mathematical modeling of infectious disease or related social diffusion processes.
```{r, eval=FALSE, tidy=FALSE}
help(package = "EpiModel")
?EpiModel
```

You may also review the documentation of the related `Statnet` packages.
```{r}
help(package = "ergm")
help(package = "tergm")
help(package = "networkDynamic")
```

<br>
<hr style="background-color:#909090;height:1px;width:100%">
<small>
*Last updated:* `r format(Sys.time(), "%Y-%m-%d")`
</small>
